Local Path Planning with Moving Obstacle Avoidance based on Adaptive MPC in ATLASCAR2

### 2019-04-19 Thesis_Latex/ThesisFinal/Chapter3/Chapter3.tex
update final thesis

\section{Adaptive Model Predictive Control}
We understood that Model Predictive Control is an advanced method that predicts future behavior using a linear-time-invariant (LTI) dynamic model. These predictions are not exact and a good strategy is to make MPC insensitive to prediction errors. If the plant is strongly nonlinear or its characteristics vary dramatically with time, MPC performance might become unacceptable because LTI prediction accuracy degrade \cite{mpctoolbox}. A method that can address this degradation by adapting the prediction model for changing operating conditions is called Adaptive MPC: this control strategy uses a fixed model structure, but allows the model parameters to evolve with time. Ideally, whenever the controller requires a prediction, it uses a model appropriate for the current conditions. At each control interval, the adaptive MPC controller updates the plant model and nominal conditions. Once updated, the model and conditions remain constant over the prediction horizon. The plant model used as the basis for the adaptive MPC must be an LTI discrete-time, state-space model with a structure as follows:
\begin{equation}
\label{eqn:Adaptive_MPC_plant_discrete}
\begin{aligned}
\vec{x}(k+1)&=\vec{A}\vec{x}(k)+ \vec{B}_u \vec{u}(k)+\vec{B}_v \vec{v}(k)+\vec{B}_d \vec{d}(k)\\
\vec{y}(k)&=\vec{C}\vec{x}(k) + \vec{D}_v \vec{v}(k)+ \vec{D}_d \vec{d}(k)
\end{aligned}
\end{equation}
where the matrices \vec{A}, $\vec{B}_u$, $\vec{B}_v$, $\vec{B}_d$, \vec{C}, $\vec{D}_v$ and $\vec{D}_d$ can vary with time. The other parameters in the previous expression (\ref{eqn:Adaptive_MPC_plant_discrete}) are:
\begin{itemize}
	\item $k$ is the time index/current control interval;
	\item \vec{x} are the plant model states;
	\item \vec{u} are the manipulated inputs that can be adjusted by the MPC controller;
	\item \vec{v} are the measured disturbance inputs;
	\item \vec{d} are the unmeasured disturbance inputs;
	\item \vec{y} are the plant outputs, including both measured (necessary at least one) and unmeasured.
\end{itemize}
In the adaptive MPC control, there are additional requirements for the plant model, like the sample time $T_s$ that has to be constant and identical to the MPC control interval. This control strategy prohibits direct feed-through from any manipulated variable to any plant output. Thus, $\vec{D_v} = \vec{0}$ in the above model.
A traditional MPC controller includes a nominal operating point at which the plant model applies, such as the condition at which you linearize a nonlinear model to obtain the LTI approximation (equilibrium, reference trajectory and the	
most updated value) \cite{mpctoolbox}. In adaptive MPC, as time evolves it should update the nominal operating point to be consistent with the updated plant model. It is possible to rewrite the plant model in terms of deviations from the nominal conditions as follows:
\begin{equation}
\label{eqn:Adaptive_MPC_nominal_condition}
\begin{aligned}
\vec{x}(k+1)&=\overline{\vec{x}}+\vec{A}(\vec{x}(k)-\overline{\vec{x}})+ \vec{B}(\vec{u}_t(k)-\overline{\vec{u}}_t)+\overline{\Delta \vec{x}} \\
\vec{y}(k)&=\overline{\vec{y}}+\vec{C}(\vec{x}(k)-\overline{\vec{x}}) + \vec{D}(\vec{u}_t(k)-\overline{\vec{u}}_t)
\end{aligned}
\end{equation}
where the matrices \vec{A}, \vec{B}, \vec{C} and \vec{D} are updated with respect to time. The other parameters in the previous structure (\ref{eqn:Adaptive_MPC_nominal_condition}) are:
\begin{itemize}
	\item $\vec{u}_t$ is the combined plant input variable, comprising $\vec{u}$, $\vec{v}$ and $\vec{d}$ variables defined earlier;
	\item $\overline{\vec{x}}$ are the nominal states;
	\item $\overline{\Delta \vec{x}}$ are the nominal state increments;
	\item $\overline{\vec{u}_t}$ and $\overline{\vec{y}}$ are the nominal inputs and outputs.
\end{itemize} 
The adaptive MPC uses a Kalman filter to update its controller states which include the plant, the disturbance and measurement noise model states. In particular this filter is linear-time-varying (LTV) because adjusts the gains at each control
interval to maintain consistency with the updated plant model.

>>>>

\section{Adaptive Model Predictive Control}
We understood that Model Predictive Control is an advanced method that predicts future behavior using a linear-time-invariant (LTI) dynamic model. These predictions are not exact and a good strategy is to make MPC insensitive to prediction errors. If the plant is strongly nonlinear or its characteristics vary dramatically with time, MPC performance might become unacceptable because LTI prediction accuracy degrade \cite{mpctoolbox}.

A method that can address this degradation by adapting the prediction model for changing operating conditions is called Adaptive MPC: this control strategy uses a fixed model structure, but allows the model parameters to evolve with time. Ideally, whenever the controller requires a prediction, it uses a model appropriate for the current conditions. At each control interval, the adaptive MPC controller updates the plant model and nominal conditions. Once updated, the model and conditions remain constant over the prediction horizon. The plant model used as the basis for the adaptive MPC must be an LTI discrete-time, state-space model with a structure as follows:
\begin{equation}
\label{eqn:Adaptive_MPC_plant_discrete}
\begin{aligned}
\vec{x}(k+1)&=\vec{A}\vec{x}(k)+ \vec{B}_u \vec{u}(k)+\vec{B}_v \vec{v}(k)+\vec{B}_d \vec{d}(k)\\
\vec{y}(k)&=\vec{C}\vec{x}(k) + \vec{D}_v \vec{v}(k)+ \vec{D}_d \vec{d}(k)
\end{aligned}
\end{equation}
where the matrices \vec{A}, $\vec{B}_u$, $\vec{B}_v$, $\vec{B}_d$, \vec{C}, $\vec{D}_v$ and $\vec{D}_d$ can vary with time. The other parameters in the previous expression (\ref{eqn:Adaptive_MPC_plant_discrete}) are:
\begin{itemize}
	\item $k$ is the time index/current control interval;
	\item \vec{x} are the plant model states;
	\item \vec{u} are the manipulated inputs that can be adjusted by the MPC controller;
	\item \vec{v} are the measured disturbance inputs;
	\item \vec{d} are the unmeasured disturbance inputs;
	\item \vec{y} are the plant outputs, including both measured (necessary at least one) and unmeasured.
\end{itemize}
In the adaptive MPC control, there are additional requirements for the plant model, like the sample time $T_s$ that has to be constant and identical to the MPC control interval. This control strategy prohibits direct feed-through from any manipulated variable to any plant output. Thus, $\vec{D_v} = \vec{0}$ in the above model.
A traditional MPC controller includes a nominal operating point at which the plant model applies, such as the condition at which you linearize a nonlinear model to obtain the LTI approximation (equilibrium, reference trajectory and the	
most updated value) \cite{mpctoolbox}.

In adaptive MPC, as time evolves it should update the nominal operating point to be consistent with the updated plant model. It is possible to rewrite the plant model in terms of deviations from the nominal conditions as follows:
\begin{equation}
\label{eqn:Adaptive_MPC_nominal_condition}
\begin{aligned}
\vec{x}(k+1)&=\overline{\vec{x}}+\vec{A}(\vec{x}(k)-\overline{\vec{x}})+ \vec{B}(\vec{u}_t(k)-\overline{\vec{u}}_t)+\overline{\Delta \vec{x}} \\
\vec{y}(k)&=\overline{\vec{y}}+\vec{C}(\vec{x}(k)-\overline{\vec{x}}) + \vec{D}(\vec{u}_t(k)-\overline{\vec{u}}_t)
\end{aligned}
\end{equation}
where the matrices \vec{A}, \vec{B}, \vec{C} and \vec{D} are updated with respect to time. The other parameters in the previous structure (\ref{eqn:Adaptive_MPC_nominal_condition}) are:
\begin{itemize}
	\item $\vec{u}_t$ is the combined plant input variable, comprising $\vec{u}$, $\vec{v}$ and $\vec{d}$ variables defined earlier;
	\item $\overline{\vec{x}}$ are the nominal states;
	\item $\overline{\Delta \vec{x}}$ are the nominal state increments;
	\item $\overline{\vec{u}_t}$ and $\overline{\vec{y}}$ are the nominal inputs and outputs.
\end{itemize} 
The adaptive MPC uses a Kalman filter to update its controller states which include the plant, the disturbance and measurement noise model states. In particular this filter is linear-time-varying (LTV) because adjusts the gains at each control
interval to maintain consistency with the updated plant model.

### 2019-04-19 Thesis_Presentation/Presentazione/Presentazione.tex
update final thesis

\documentclass{beamer}
\usepackage[italian]{babel}
\usepackage[utf8]{inputenc}
\usepackage{default}
\usepackage{afterpage}
\usepackage{xcolor}
\usepackage{wrapfig}
\usepackage{caption,subcaption}
\usepackage[font={small,it}]{caption}
\usepackage[font={scriptsize,it}]{subcaption}
\usepackage{bbm}
\usepackage{siunitx}
%\usepackage{movie15}
%\usepackage{multimedia} 
%\newsavebox{\tempbox}
\renewcommand{\vec}[1]{\ensuremath{\boldsymbol{\mathit{#1}}}}
\usetheme{Padova}
\setbeamertemplate{caption}[numbered]
\title{Local Path Planning with Moving Obstacle Avoidance based on Adaptive MPC in ATLASCAR2}
\subtitle{\itshape Tesi di Laurea in Ingegneria dell'Automazione}
\author{Relatore: Prof. Angelo Cenedese\qquad\quad Laureando: Alberto Franco Correlatore: Prof. Vitor Santos}
\date{\vspace{0.4cm}15 Aprile 2019}

@@

	\begin{frame}{Model Predictive Control}
		Model Predictive Control:
		\begin{itemize}
			\item tecnica avanzata basata su ottimizzazione matematica
			\item prevede il comportamento futuro utilizzando un modello dinamico LTI
			\item predizioni non esatte - insensibile agli errori di predizione - performance inaccettabili
		\end{itemize}
		$\Longrightarrow$ Tecnica adottata: {\bfseries Adaptive Model Predictive Control}
		\begin{itemize}
			\item adatta il modello di predizione per cambiare le condizioni operative
			\item struttura di modello fissa, ma consente ai parametri del modello di evolvere nel tempo
		\end{itemize}
	\end{frame}

@@

	\begin{frame}{Conclusioni e Sviluppi Futuri}
		\begin{columns}[onlytextwidth,T]
			\begin{column}[c]{0.75\textwidth}
				\textbf{Conclusioni}:			\begin{itemize}
					\item Sistema di anticollisione in ambiente dinamico (ostacoli in movimento)
					\item Sistema di assistenza al mantenimento della corsia (Lane Keeping Assist)
				\end{itemize}
				\textbf{Sviluppi futuri}:
				\begin{itemize}
					\item Combinazione dei due metodi in un unico schema di controllo
					\item Implementazione in ambiente ROS-Gazebo
					\item Uso di dati reali collezionati dai sensori dell'ATLASCAR2 per testare gli algoritmi sviluppati
				\end{itemize}
			\end{column}
			\begin{column}[c]{0.20\textwidth}
				\begin{figure}
					\vspace{-3em}
					\includegraphics[width=
					\textwidth]{./images/GAZEBO.png}
				\end{figure}
				\begin{figure}
					\vspace{-3em}
					\includegraphics[width=
					\textwidth]{./images/ROS.png}
				\end{figure}
			\end{column}			
		\end{columns}
	\end{frame}

\end{document}

>>>>

\documentclass{beamer}
\usepackage[italian]{babel}
\usepackage[utf8]{inputenc}
\usepackage{default}
\usepackage{afterpage}
\usepackage{xcolor}
\usepackage{wrapfig}
\usepackage{caption,subcaption}
\usepackage[font={small,it}]{caption}
\usepackage[font={scriptsize,it}]{subcaption}
\usepackage{bbm}
\usepackage{siunitx}
\usepackage{amsmath,amssymb}
%\usepackage{movie15}
%\usepackage{multimedia} 
%\newsavebox{\tempbox}
\renewcommand{\vec}[1]{\ensuremath{\boldsymbol{\mathit{#1}}}}
\usetheme{Padova}
\setbeamertemplate{caption}[numbered]
\title{Local Path Planning with Moving Obstacle Avoidance based on Adaptive MPC in ATLASCAR2}
\subtitle{\itshape Tesi di Laurea in Ingegneria dell'Automazione}
\author{Relatore: Prof. Angelo Cenedese\qquad\quad Laureando: Alberto Franco Correlatore: Prof. Vitor Santos}
\date{\vspace{0.4cm}15 Aprile 2019}

@@

	\begin{frame}{Model Predictive Control - parte 1}
		Model Predictive Control:
		\begin{itemize}
			\item tecnica avanzata basata su ottimizzazione matematica
			\item prevede il comportamento futuro utilizzando un modello dinamico LTI
			\item predizioni non esatte - insensibile agli errori di predizione - performance inaccettabili
		\end{itemize}
		$\Longrightarrow$ Tecnica adottata: {\bfseries Adaptive Model Predictive Control}
		\begin{itemize}
			\item adatta il modello di predizione per cambiare le condizioni operative
			\item struttura di modello fissa, ma consente ai parametri del modello di evolvere nel tempo
		\end{itemize}
	\end{frame}

	\begin{frame}{Model Predictive Control - parte 2}
		Dato un sistema discreto LTI:
		\begin{equation*}
		\label{eqn:MPC_plant_discrete}
		\vec{x}(k+1)=\vec{A}\vec{x}(k)+ \vec{B} \vec{u}(k)
		\end{equation*}
		%dove $\vec{x}(k)\in\mathbb{R}^n$, $\vec{u}(k)\in\mathbb{R}^m$.
		L'idea è quella di minimizzare la funzione obiettivo assicurando che i vincoli siano rispettati:
		\begin{equation*}
		\label{eqn:MPC_optimization}
		\begin{aligned}
		& \underset{\textbf{u}}{\text{minimizzare}}
		& & {\text{\itshape
				\bfseries J}}(\vec{x}(k), \textbf{u}) \\
		& \text{soggetto a}
		& & \vec{x}_{k+i+1} = \vec{A}\vec{x}_{k+i}+ \vec{B} \vec{u}_{k+i}\quad\forall i=0,\dots N-1;\\
		& & & \vec{x}_{k+i}\in \mathcal{X}\quad\forall i=0,\dots N-1;\\
		& & & \vec{u}_{k+i}\in \mathcal{U}\quad\forall i=0,\dots N-1;\\  
		& & & \vec{x}_{k+N}\in \mathcal{X}_f;\quad\vec{x}_k = \vec{x}(k).
		\end{aligned}
		\end{equation*}
		%dove $\textbf{u}=(\vec{u}_k,\dots,\vec{u}_{k+N-1})$ sono una sequenza di ingressi di controllo, $\vec{x}_{k+i}$ è lo stato al tempo $k+i$ come predetto al tempo $k$, $N$ è l'orizzonte di predizione. Gli insiemi $\mathbb{X}\in\mathbb{R}^n$ e $\mathbb{U}\in\mathbb{R}^m$ definiscono i vincoli sullo stato e sull'ingresso.Infine l'insieme $\mathbb{X}_f\subseteq\mathbb{X}$ definisce il vincolo finale sullo stato. 
	\end{frame}

@@

	\begin{frame}{Conclusioni e Sviluppi Futuri}
		\begin{columns}[onlytextwidth,T]
			\begin{column}[c]{0.79\textwidth}
				\textbf{Conclusioni}:			\begin{itemize}
					\item Sistema di anticollisione in ambiente dinamico
					\item Sistema Lane Keeping Assist
				\end{itemize}
				\textbf{Sviluppi futuri}:
				\begin{itemize}
					\item Combinazione in un unico schema di controllo
					\item Implementazione in ambiente ROS-Gazebo
					\item Dati reali collezionati dai sensori dell'ATLASCAR2 per testare gli algoritmi
				\end{itemize}
			\end{column}
			\begin{column}[c]{0.20\textwidth}
				\begin{figure}
					\vspace{-3em}
					\includegraphics[width=
					\textwidth]{./images/GAZEBO.png}
				\end{figure}
				\begin{figure}
					\vspace{-3em}
					\includegraphics[width=
					\textwidth]{./images/ROS.png}
				\end{figure}
			\end{column}			
		\end{columns}
		\vspace{1em}
		$\Longrightarrow$ A. Franco, 
		V. Santos, {\itshape ``Short-term Path Planning with Multiple Moving Obstacle Avoidance based on Adaptive MPC''}, IEEE ICARSC 2019, Porto, April 24-26
	\end{frame}

	\section{Grazie per l'attenzione! \qquad\qquad{\itshape\large(Obrigado)}}
\end{document}
